---
title: 'Puzzle Activity: Writing Functions'
author: "Nathan Diekema"
date: "10/5/2021"
output: 
  rmdformats::downcute:
    lightbox: true
    self_contained: true
    gallery: true
    highlight: github
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```


#### **1. Write a function called divide_and_round(), which takes a number or vector of numbers, divides them all by their smallest value, and rounds the results to the nearest whole number.**


```{r message=FALSE, warning=FALSE}

divide_and_round <- function(nums){
  
  my_min <- min(nums, na.rm = TRUE)
  divided <- nums / my_min
  return (round(divided))
}

test <- c(1:5, NA)

divide_and_round(test)

```


#### **2. Write a function called no_nines_or_twelves(), which takes a number or vector of numbers and returns TRUE if the number is NOT divisible by 9 or 12, FALSE otherwise.**


```{r, message=FALSE, warning=FALSE}
no_nines_or_twelves <- function(nums){
  check <- !(nums %% 9 == 0 | nums %% 12 == 0)
  return(check)
}

test <- c(1:10*5, NA)

no_nines_or_twelves(test)


```


#### **3. Write a function called every_other() which takes a vector and returns every other value in the vector. Accomplish this with a for loop. (Although there are certainly better ways to do it!)**


```{r, message=FALSE, warning=FALSE}

every_other <- function(vec, start=1){
  k <- vector()
  count = 0
  
  for (i in start:length(vec)) {
    if (count %% 2 == 0) {
      k = c(k, vec[i])
    }
    count = count + 1
  }
  
  return(k)
}

test <- c(1:10)

every_other(test)
every_other(test, start = 2)

```


#### **4. Write a function called shorten() which takes a vector and keeps dropping the first value, until the sum of the vector is less than 350 Return the remaining values.**


```{r}
shorten <- function(vec) {
  sum_vec = sum(vec, na.rm=TRUE)
  while (sum_vec > 350) {
    vec = vec[-1]
    sum_vec = sum(vec, na.rm=TRUE)
  }
  return(vec)
}

test <- c(1:100, NA)
shorten(test)

test <- c(NA, 1:50)
shorten(test)
```


#### **5. Put it all together**


```{r}
my_vec <- c(39, 1.87, 48, 11, 8, 45, 21, 5, 12, 33, 9, 11, 108, 4, 18, 5, 16, 17, 8, 48, 27, 24, 4, 22, 12, 27, 23, 46, 42, 35, 15, 34, 36, 26, 18, 10, 18.21, 72.04, 36.9, 41.81, 29, 89.75, 34.03, 20.18, 48.74, 15.76, 31.86, 83.6, 43.55, 39.99, 23.55, 8.54, 24.71, 22.02, 9.71, 62.14, 35.46, 16.61, 15.66, 21.29, 30.52, 201.07, 45.81, 7.85, 30.13, 34.14, 22.62, 10.2, 6.02, 30.12, 10.64, 31.72, 24.57, 14.43, 43.37, 89.93, 44.72, 51.32, 13.62, 45.56, 22.96, 7.05, 29.99, 41.38, 26.59, 23.04, 19.82, 50.73, 39.56, 43.79, 30.22, 85.85, 5.78, 78.85, 29.52, 66.27, 44.06, 27.28, 24.43, 64.32, 3.35, 67.45, 46.72, 48.44, 48.65, 33.3, 40.28, 19.04)

my_vec <- every_other(my_vec, start = 2)
my_vec <- divide_and_round(my_vec)
my_vec <- every_other(my_vec, start = 1)
my_vec <- shorten(my_vec)
my_vec <- my_vec[no_nines_or_twelves(my_vec)]

my_vec <- sort(my_vec)

my_vec

```





