---
title: 'Activity 7: Character Data & Regular Expressions'
author: "Nathan Diekema"
date: "10/12/2021"
output: 
  rmdformats::downcute:
    lightbox: true
    self_contained: true
    gallery: true
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center", fig.height=5, fig.width=10)
```

## Introduction

The NBA Player of the Week dataset (NBA_player_of_the_week.csv) on Canvas contains information on the player of the week from the 1984-1985 season all the way up through the 2017-2018 season.\


```{r, message=FALSE, warning=FALSE}
library(tidyverse)

nba_data <- read_csv("data/NBA_player_of_the_week.csv")
head(nba_data)
```

## Exercises

### Step 1

For each variable identify its type. Which variables are NOT numeric? Does this make sense?\

```{r, message=FALSE, warning=FALSE}
spec(nba_data)
```

**Non-numeric Variables**

- Conference
- Date
- Height
- Player
- Position
- Season
- Team
- Weight

A few of these columns could easily be recorded as numeric variables, but they're not. Especially variables such as height and weight.


### Step 2

Notice that the last 100 or so rows of the dataset have height values in centimeters. The other height values have a specific format, “feet-inches”. Create a new variable called “heightcm” that contains the height of each player in centimeters. This variable should be numeric.\

```{r, message=FALSE, warning=FALSE}

nba_data <- nba_data %>% 
  mutate(heightcm = lapply((str_extract_all(Height, "[:digit:]+")), as.numeric)) %>% 
  mutate(heightcm = ifelse(lengths(heightcm) == 2,
                           lapply(heightcm, function(x) round(x[1]*30.48+x[2]*2.54)),
                           heightcm)) %>% 
  mutate(heightcm = do.call(rbind, lapply(heightcm, as.numeric)))

head(nba_data$heightcm)

nba_data

```

### Step 3

Notice that the last 100 or so rows of the dataset have weight values in kilograms. The rest of the values are in lbs. Create a new variable called “weightkg” that contains the weight of each player in kilograms. This variable should be numeric.\


```{r message=FALSE, warning=FALSE}

nba_data <- nba_data %>% 
  mutate(weightkg = ifelse(str_detect(Weight, "kg"), 
                           as.numeric(str_extract(Weight, "[:digit:]+")), 
                           as.numeric(str_extract(Weight, "[:digit:]+"))*0.45359)) %>% 
  mutate(weightkg = round(weightkg))
         
matrix(head(nba_data$weightkg))

```


### Step 4

Create a new variable called “timesWon” that contains the number of times each Player/Team combination won. Your resulting dataset should only have one row for each Player/Team combination. Note: we will lost the date-related variables with this, but it’s fine!\


```{r message=FALSE, warning=FALSE}

comboWins <- nba_data %>% 
  select(Player, Team) %>% 
  mutate(
    player_team = str_c(Player, Team, sep=" | ")
  ) %>% 
  group_by(player_team, Player, Team) %>% 
  summarize(timesWon = n())

comboWins %>% 
  arrange(desc(timesWon))

# Merge into nba dataset
df1 <- data.frame(nba_data)
df2 <- data.frame(comboWins)
nba_data <- merge(df1, df2[,c("Player","Team","timesWon")])
  
```


### Step 5

Which 5 players have won player of the week the most times?

```{r message=FALSE, warning=FALSE}
library(kableExtra)

comboWins <- comboWins %>% 
  arrange(desc(timesWon))

head(comboWins[,c("Player","Team","timesWon")], 5) %>% 
  kbl(caption="Top 5 players who have won player of the week") %>% 
  kable_styling(full_width=F, position="center")

```


### Step 6

Fill in the blanks in the following paragraph after writing code to do the relevant calculations.

```{r message=FALSE, warning=FALSE}
# Num rows
nrow(nba_data)

# Num positions
n_distinct(nba_data$Position)

# Age range
range(nba_data$Age)

# % From SG position
(nba_data %>% 
  filter(Position=="SG") %>% 
  distinct(Player) %>% 
  count()) / n_distinct(nba_data$Player)

# Median height (cm)
median(nba_data$weightkg, rm.NA=T)

# Most & least common team
nba_data %>% 
  group_by(Team) %>% 
  count() %>% 
  arrange(desc(n))

# Lakers Percentage won
71 / nrow(nba_data)

# Max weight in eastern conference
nba_data %>% 
  filter(Conference=="East") %>% 
  summarize(max_weight=max(weightkg))

# % of players who have won > 10 games
nba_data %>% 
  filter(timesWon > 10) %>% 
  group_by(Player) %>% 
  count() %>% 
  nrow() / n_distinct(nba_data$Player)

# % of players who have won > 10 games - East Conference
nba_data %>% 
  filter(Conference == "East", timesWon > 10) %>% 
  group_by(Player) %>% 
  count() %>% 
  nrow() / n_distinct(nba_data$Player)

# % of players who have won > 10 games - West Conference
nba_data %>% 
  filter(Conference == "West", timesWon > 10) %>% 
  group_by(Player) %>% 
  count() %>% 
  nrow() / n_distinct(nba_data$Player)

```

This data set contains **1145** rows containing information on **11** different positions. Players’ age ranged from **19** to **40**, **12.4%** are from the SG position, and the overall average Height value (in centimeters) is **200.9**. The most common team is **Los Angeles Lakers** (**6.2%**), while the least common team was **New Orleans Pelicans**. Although the overall median Weight (in kilograms) was **102**, the maximum Weight among East conference players was **147kg**. Overall, **7.299%** percent of these players have won more than 10 times; among the West conference this was **1.824%** and among the East conference this was **3.649%**.



