---
title: "Lab 2 Challenge"
author: "Nathan Diekema"
date: "10/4/2021"
output: 
  prettydoc::html_pretty: 
    theme: leonids
    fig_retina:	2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center", fig.height=5, fig.width=10)
```

## Load in and clean data

```{r message=FALSE, warning=FALSE}
library(tidyverse)

hiphop <- read_csv("data/hiphop.csv")
```

```{r message=FALSE, warning=FALSE}

hiphop_clean <- 
  hiphop %>% 
  drop_na(c(ethnic, sex, age)) %>% 
  mutate_at(vars(ethnic, fam1), as.factor)


hiphop_clean
```


**Words that are recognizeable to those who describe themselves as hiphop fans, but less recognizeable to those who do not consider themselves fans**

```{r message=FALSE, warning=FALSE}

## Find words that are recognizable by "fans" of hiphop
words1 <- 
  hiphop_clean %>% 
  select(word, familiarity, hiphop) %>% 
  filter(hiphop > 3) %>% 
  group_by(word) %>% 
  summarize(
    avg_fam_fans=mean(familiarity)
  )

## Find words that are recognizable by "fans" of hiphop
words2 <- 
  hiphop_clean %>% 
  select(word, familiarity, hiphop) %>% 
  filter(hiphop < 2) %>% 
  group_by(word) %>% 
  summarize(
    avg_fam_not_fans=mean(familiarity)
  )

## Combine the two tables
compare_words <- left_join(x=words1, y=words2, by="word")

## Add a column with the difference between familiarity
compare_words %>% 
  mutate(difference = abs(avg_fam_fans - avg_fam_not_fans)) %>% 
  arrange(desc(difference))

```

**Words ranked based on familiarity with people that live in the city and who are hiphop fans**

```{r message=FALSE, warning=FALSE}
popular_words <-
  hiphop_clean %>% 
  filter(city >= 50000,
        hiphop > 3) %>% 
  group_by(word) %>% 
  summarize(
    avg_fam=mean(familiarity)
  )

popular_words %>% 
  arrange(desc(avg_fam))

```


**Hella Bling Track List**

1. Finna get loose 
2. What it do (feat. Nicki Minaj)
3. Make it rain 
4. Boo thang
5. You feel me?
6. Make it rain
7. Ballin'
8. A-town girl
9. You know what it is
10. Hella chedda
11. Player hater

For the track names on this album I used some of the words with the highest average familiarity for people who live in decently large cities and who likely consider themselves fans of hiphip. I picked these filters because I figured the people who are most likely going to listen to this album in the first place are those who are fans of hiphop and also the majority of the population lives in cities. I incorporated many of the most familiar words in the track names to appeal to people in this population.

